---
title: "R Project on Graduates Placement"
author: "Jenita Jeyakumar"

output: html_document
---

### SECTION-1 INTRODUCTION:

The data set used for this project is from Kaggle.com. It contains information about the placement status of graduates from an MBA course. The variables include the percentage, board and/or stream of secondary school, higher secondary school, degree, and MBA, of male and female students. The data also contains Employability test scores and work experience status for all students and the salary of the placed students.

#### [RESEARCH QUESTION:]{.smallcaps}

As part of the main analysis of this data set, we will look into factors that determine whether or not a student gets placed.

[**Data source:**]{.smallcaps} <https://www.kaggle.com/benroshan/factors-affecting-campus-placement>

```{r, echo = FALSE, message=FALSE}
library(tidyverse)
library(readr)
library(dplyr)
```

```{r, echo = FALSE, message=FALSE}
#IMPORTING THE DATASET
placement <- read_csv("C:/Users/Jenita Jeyakumar/Desktop/DAB501/Data/Placement_Data_Full_Class.csv")
```
Taking a look at the data using the "**glimpse**" function:

```{r, echo = TRUE}
glimpse(placement)
```

### SECTION-2 DATA ANALYSIS PLAN:

-   The outcome or response variable (Y) is the "status" which is a categorical variable that gives info about the status of placement of the students. It has two values, either placed or not placed. It can also be treated as a binary variable since it has two values- placed and not placed.

    The explanatory or predictor variables (X) could be any of the following :

    1.  SSC or senior secondary school percentage

    2.  HSC or higher secondary school percentage

    3.  Degree percentage and specialization

    4.  Employability test percentage

    5.  MBA percentage and specialization

    6.  Work experience

-   The entire data set can be grouped using any of the following variables:

    1.  Gender (Female and Male)

    2.  Work experience (Yes and No)

    3.  MBA specialization (Mkt&Fin and Mkt&HR)

-   The preliminary analysis will include some basic univariate analysis.

    Below gives the summary statistics for the dataset:

    ```{r, echo = TRUE}
    summary(placement)
    ```

    #### PRELIMINARY ANALYSIS:

    -   Below is the distribution of students by Degree stream:

    ```{r, echo = FALSE}
    ggplot(placement, aes(degree_t)) + geom_bar(fill = "red") + labs(title = "Distribution by Degree stream", x = "Degree Stream", y = "Count")
    ```

    -   Below is the distribution of students by MBA specialization:

    ```{r, echo = FALSE}
    ggplot(placement, aes(specialisation)) + geom_bar(fill = "grey") + labs(title = "Distribution by MBA specialization", x = "MBA specialization", y = "Count")
    ```

    -   Below is the distribution of students by Gender:

    ```{r, echo = FALSE}
    ggplot(placement, aes(gender)) + geom_bar(fill = "blue") + labs(title = "Distribution by Gender", x = "Gender", y = "Count")
    ```

-   We will determine if Data cleaning and transformation steps are required and the same will be executed:

    -   The summary shows that there are NA values in the 'salary' variable. These NA's correspond to the salary for students that are not placed, hence not making it necessary to clean the NA values because it makes sense to have them.

-   We will check if there are outliers for the variable 'salary' in the data using a boxplot, and will see how to handle them.

-   The statistical method 'Linear regression' can be used to determine relationship between various factors that lead to placement.

# **PROJECT- ANALYSIS and VISUALIZATIONS**

#### 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## **A. DATA CLEANING AND TRANSFORMATION:**

##### 

-   The summary of the data set shows that there are 67 NA values in the 'salary' variable. These NA's correspond to the salary for students that are not placed, hence not making it necessary to clean the NA values because it makes sense to have them.

-   We will check if there are outliers for the variable 'salary' in the data using a boxplot, and will see how to handle them.

    ```{r, echo=FALSE}
    ggplot(placement, aes(salary), na.rm = TRUE) + geom_boxplot() + labs(title = "Distribution of Salary of placed students", x = "Salary")
    ```

    Let's check the observations corresponding to the top 5 salary:

    ```{r, echo=FALSE}
    p1 <- placement %>% arrange(desc(salary))
    head(p1, 5)
    ```

    The high values of salary are justified for these observations as the percentages are high for these students, and the top 4 students have work experience too.

-   The 'status' variable says whether or not the student is placed. We can use the mutate function to make a new column to obtain a numeric representation of the 'status' variable, where the value **'not placed' is 1 and 'placed' is 2.**

    ```{r}
    placement <- placement %>% mutate(status_int = as.numeric(factor(status)))
    placement
    ```

#### 

## **B. ANALYSIS QUESTIONS:**

### **B.1. What is the placement status of the top 10 students in the employability test:**

We can see that 20% of the students are not placed in the top 10 performers of the employability test. Indicating that the test is not the only determining factor for the students' placement.

```{r, echo=FALSE}
p1 <- placement %>% arrange(desc(etest_p)) %>% select(-sl_no, -ssc_b, -ssc_p, -hsc_p, -hsc_b, -hsc_s, -degree_t, -specialisation)
p2 <- head(p1,10)
p2
```

```{r}
ggplot(p2, aes(status)) + geom_bar(fill = "pink") + labs(title = "Placement status of top 10 scorers in Employability test", x = "Placement status", y = "Count")
```

To further prove the above inference let us plot the distribution of the **Employability test scores for all students** as follows:

As we can observe from the below plot, the peak for the students not placed is of lower value but that of students placed has a bimodal distribution, with lower and higher peaks of test scores. This indicates that the Employability test scores is NOT the only determining factor for the placement of students.

```{r, echo=FALSE}
ggplot(placement, aes(x=etest_p)) + 
  geom_density(aes(fill =status), alpha = 0.5, adjust = 1) + 
  labs(title = "Employability test score distribution by Placement Status", x = "Employability test score", y = "Count", fill = "Placement Status")
```

#### 

### **B.2. Which stream of Degree/UG has a larger proportion of students not placed?**

The students in Other streams of Degree/UG apart from Comm&Mgmt and Sci&Tech have faced a larger proportion of unemployment after completing their PG studies. While the proportion of students placed from the streams Comm&Mgmt and Sci&Tech is the same at around 75%.

```{r, echo=FALSE}
ggplot(placement, aes(degree_t)) + 
  geom_bar(aes(fill=status), position = "fill") + 
  labs(title = "Distribution of Degree Type by Placement Status", x = "Degree Type", y = "Percentage Distribution", fill = "Placement status")
```

#### 

### **B.3. Do higher percentages in SSC and HSC help in getting placement offers?**

As we can observe from the scatter plot below, the students with lower percentages are not placed whereas, the ones with higher scores in SSC and HSC have gotten placed.

```{r, echo=FALSE}
ggplot(placement, aes(x=ssc_p, y=hsc_p)) + 
  geom_point(aes(color=status)) + 
  geom_smooth(se = TRUE) + 
  labs(title = "SSC vs HSC Percentage distribution by Placement Status", x = "SSC Percentage", y = "HSC Percentage", colour = "Placement Status")
```

#### 

### **B.4. Does the percentage obtained in Degree impact the placement status?**

As we can observe from the below plot, the peak for the students placed is distinctly farther from that of students not placed. This indicates that the Degree percentage was a major deciding factor for the placement of the students.

```{r, echo=FALSE}
ggplot(placement, aes(x=degree_p)) + 
  geom_density(aes(fill =status), alpha = 0.5, adjust = 1.5) + 
  labs(title = "Degree Percentage distribution by Placement Status", x = "Degree Percentage", y = "Count", fill = "Placement Status")
```

#### 

### **B.5. Does the percentage obtained in MBA impact the placement status?**

As we can observe from the below plot, the peak for the students placed is the same as that of students not placed. This indicates that the MBA percentage was NOT a major deciding factor for the placement of the students.

```{r, echo=FALSE}
ggplot(placement, aes(x=mba_p)) + 
  geom_density(aes(fill =status), alpha = 0.5, adjust = 1.5) + 
  labs(title = "MBA Percentage distribution by Placement Status", x = "MBA Percentage", y = "Count", fill = "Placement Status")

```

#### 

### **B.6. From the above points, we can see that the SSC, HSC and Degree percentages are impacting the placements. Let us see if the average value has the same behavior as each individual entity.**

The below graph proves that the behavior of the average is the same as previously mentioned individual entities like SSC, HSC and Degree percentage. Higher the average percentage, better are the placement records.

```{r}
placement <- placement %>% mutate(avg_ug = (ssc_p + hsc_p + degree_p)/3)
placement
```

```{r, echo=FALSE}
ggplot(placement, aes(x=avg_ug)) + 
  geom_density(aes(fill =status), alpha = 0.5, adjust = 1.5) + 
  labs(title = "Average Percentage Distribution until UG by Placement Status", x = "Average Percentage", y = "Count", fill = "Placement Status")
```

#### 

### **B.7. Does having work experience boost one's chances of getting placed?**

Yes, having work experience has bettered the chances of students to get job offers. As we can see from the 100% stacked bar graph, the proportion of students placed is higher for those having work experience.

```{r, echo=FALSE}
ggplot(placement, aes(x=workex)) + 
  geom_bar(aes(fill =status), alpha = 0.5, position = "fill") + 
  labs(title = "Work experience distribution by Placement Status", x = "Work experience", y = "Percentage distribution", fill = "Placement Status")

```

#### 

### **B.8. Is there gender bias in the pay for students who are placed?**

From the graph below, we can observe that the pay for Female students is slightly lesser than that of the Male counterparts, as the peaks are at a lower Salary value for the females.

```{r, echo=FALSE}
p3 <- placement %>% filter(salary != "NA")
ggplot(p3, aes(salary)) + 
  geom_density(aes(fill = gender), alpha = 0.5, adjust = 2) + 
  labs(title = "Salary distribution by Gender", x = "Salary of placed students", y = "Density", fill = "Gender")
```

#### 

## **C. STATISTICAL MODEL: LINEAR REGRESSION**

The Equation for the LINEAR REGRESSION between the target or predicted variable (Y) i.e. "status_int" vs the explanatory variable (X) i.e. "avg_ug" is as follows:

(predicted mean value of Y) = intercept + (slope)(mean value of X)

Let's call the lm method to generate a linear regression model that gives the intercept and slope values as coefficients:

```{r}
linear_reg <- lm(status_int ~ avg_ug, data = placement)
linear_reg
```

Hence, our equation now becomes:

status_int = -0.84761 + (0.03804)\*avg_ug

```{r}
summary(linear_reg)
```

The above summary function when called gives the residuals' IQR range, coefficient values for intercept and slope, and the Coefficient of Determination value i.e. R-square is equal to 0.4224.

In the column that was created "status_int", the value **'not placed' is 1 and 'placed' is 2.**

-   The slope can be interpreted as follows:

    For every increase in unit of average percentage (i.e. avg_ug), it is expected that the status_int would increase on an average by 0.092209 units. Thus increasing the chances of getting placed.

-   The intercept can be interpreted as follows:

    When the average percentage scored would be zero, it is expected on average that value of status_int is -0.84761 which can never be possible since percentage values are nowhere zero.\

#### 

## **D. CONCLUSION:**

The analysis above can be concluded as follows:

-   Many factors were taken into consideration to determine what impacts the placement of students in a class who are graduating with an MBA.

-   The Employability test scores is NOT the only determining factor for the placement of students.

-   The students in Other streams of Degree/UG apart from Comm&Mgmt and Sci&Tech have faced a larger proportion of unemployment.

-   While 75% of students from the streams Comm&Mgmt and Sci&Tech are placed.

-   The students with higher percentages in SSC and HSC had better placement numbers.

-   The Degree percentage was also a major deciding factor for the placement of the students.

-   Whereas, their MBA percentage was NOT a major deciding factor for the placement of the students.

-   The behavior of the average of SSC, HSC and Degree percentage is the same as the previously mentioned individual entities SSC, HSC and Degree percentage. Higher the average percentage, better are the placement records.

-   Having work experience has bettered the chances of students to get job offers.

-   Gender bias in pay is still a controversial topic in today's times. The pay for Female students is slightly lesser than that of the Male counterparts.

-   On the whole, students with good average percentage until their UG, who are from either Comm&Mgmt or Sci&Tech Degree background, and have had prior work experience have had better placement records.

#### **--Thank You --**
